---
title: "20260202_Progress_Report_XX"
author: "Xiaoni Xu"
date: "2026-02-02"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE)
```

## Overview of Recent Work

Over the past few weeks, I focused on creating the R package structure and designing the main command for the causal mediation analysis with competing risks.

Specifically, I completed the following steps:

1.  **Created the R package**: I organized the functions into a standard R package structure (`cmcrhazard`).
2.  **Designed the command**: Implemented the `cmest_pathcomprisk` function (formerly `cm_cr_gformula_hazard`).
3.  **Updated Inputs**: I aligned the function arguments with `CMAverse` conventions (e.g., changing `L` to `D`, `M` to `mreg`).

## Data Simulation

I created a self-contained simulation script to test the function. This generates data with time-varying mediators (`M_1`, `M_2`) and competing risks (`D1`, `D2`), without relying on external files.

```{r}
# Simulation parameters
n = 200
M_0 = 100; M_1_E = 1.2; M_1_C = 3.9
# ... (coefficients omitted for brevity)

# Generate covariates and mediators
E <- rnorm(n, mean=2, sd=sqrt(0.7))
C <- rbinom(n, size=1, prob=0.4)
M1 <- rnorm(n, mean = M_0 + M_1_E * E + M_1_C * C, sd=1)
M2 <- rnorm(n, mean = 46 + 0.5 * M1 + 0.9 * E - 1.3 * C, sd=1)

# Generate Survival Times (Exponential approximation)
lambda_Y <- exp(-7.4 + 0.14 * E + 0.48 * C + 0.02 * M1)
# ... (D generation simliar)

simData <- data.frame(E, C, M1, M2, Y=Y_time, D=D_time)
```

## Running the Command

I successfully ran the simulated data through the new `cmest_pathcomprisk` command.

```{r}
library(cmcrhazard)

# Define models
mreg <- list(
  lm(M_1 ~ E + C,           data = df2),
  lm(M_2 ~ E + M_1 + C,     data = df2)
)

D <- list(
  glm(D1 ~ E + M_1 + C, data = df2, family = binomial()),
  glm(D2 ~ E + M_2 + C, data = df2, family = binomial())
)

# Run the function
results <- cmest_pathcomprisk(
  D = D,
  mreg = mreg,
  mvar = c("M_1","M_2"),
  yreg = yreg,
  avar = "E",
  a = as.numeric(quantile(df2$E, 0.25)),
  astar = as.numeric(quantile(df2$E, 0.75)),
  data = df2,
  time_points = c("time.since.first.exam_1", "time.since.first.exam_2"),
  peryr = 100000,
  nboot = 20
)
```

## Results

I obtained the following estimates from the simulation run:

*   **DE (Direct Effect)**: 102.25 (-542.11, 1101.54)
*   **IEM (Indirect Effect via Mediator)**: -25.37 (-622.64, 313.49)
*   **IED (Indirect Effect via Competing Risk)**: -1.28 (-18.8, 10.27)
*   **TE (Total Effect)**: 231.45 (-466.53, 776.57)

The results show that the pipeline is working correctly. As expected with a small sample size (n=200) and low bootstrap count, the confidence intervals are wide.

## Next Steps

In the future, I plan to:
1.  Increase the sample size and number of bootstraps to obtain more stable estimates.
2.  Apply this command to Yuchen's more complex simulated dataset to further validate the method.
