---
title: "20260223_Progress_Report_XX"
author: "Xiaoni Xu"
date: "2026-02-23"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE)
```

## Overview of Recent Work

In the most recent phase of the project, I refactored the analysis pipeline to directly utilize the `cmcrhazard` R package. This transition from local function definitions to a formal package structure ensures greater reliability and easier maintenance.

Key updates include:
- _Package Integration_: Successfully migrated the script to use `cmest_pathcomprisk` from the `cmcrhazard` package.
- _Improved Power_: Increased the bootstrap samples to 1,000 and the sample size to 5,000 to obtain more stable estimates.
- _Variable Alignment_: Renamed arguments to align with package conventions (e.g., using `D` for dropout/competing risk models).

## Model Definitions

The analysis utilized the following model structures for the mediators, competing risks, and outcome:

```{r}
# Mediator models (mreg)
mreg1 <- lm(M_1 ~ E + C, data = df_main)
mreg2 <- lm(M_2 ~ E + M_1 + C, data = df_main)
mreg <- list(mreg1, mreg2)

# Dropout/Survival models (D) - Indirect Effect through Dropout death
D1_mod <- glm(D_1 ~ E + M_1 + C, data = df_main, family = binomial())
D2_mod <- glm(D_2 ~ E + M_2 + C, data = df_main, family = binomial())
D <- list(D1_mod, D2_mod)

# Outcome model (yreg - Aalen additive hazards)
yreg <- aalen(Surv(tstart, tstop, endpt) ~ const(E) + const(M) + const(C),
    data = df_process, resample.iid = 1, n.sim = 0
)
```

## Mediation Analysis Results

The analysis was performed on a simulated dataset with a sample size of $N=5,000$ and $1,000$ bootstrap iterations.

| Effect | Estimate (95% CI) |
| :--- | :--- |
| Direct Effect (DE) | 67.59 (-39.2, 164.7) |
| Indirect Effect through Mediator (IEM) | 33.07 (-52.21, 118.88) |
| Indirect Effect through Dropout death (IED) | -0.32 (-0.81, 0.17) |
| Total Effect (TE) | 99.75 (29.62, 168.63) |

## Interpretation

Based on the analysis results:
- The _Total Effect (TE)_ is positive and statistically significant, with an estimate of 99.75.
- The path-specific effects indicate a positive direction, suggesting that the mediators partially explain the effect of the treatment on the outcome.
- We also observe a small negative effect through the competing risks (_Indirect Effect through Dropout death_).

## Next Steps
- Run the analysis with a larger sample size ($N=7,500$ or $N=10,000$) to further tighten the confidence intervals for the path-specific effects.
- Continue testing the package stability with more complex simulation scenarios and real-world mimicking datasets.
