---
title: "Weekly Progress Report: Mediation Analysis & Package Validation"
author: "Xiaoni Xu"
date: "2026-03-09"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE)
```

Key updates include:
- _Bug Fix_: Identified and resolved an ordering issue in the bootstrap loop where covariate-coefficient mismatching occurred due to duplicate IDs.
- _New Feature: Exact Match Mode_: Implemented the `refit = TRUE` parameter in `cmest_pathcomprisk`, allowing for a full non-parametric bootstrap that matches the original script's logic.
- _Model Validation_: Performed a head-to-head comparison of the original re-fitting method versus the package's default parametric approach.

## Mode Comparison & Validation Results

To ensure the package behaves as expected, I compared two different bootstrap philosophies using a subset of the SHS Mimick data ($N=500, B=20$).

### 1. Parametric/Hybrid (Default)
This mode fits models once and then draws from the parameter distributions (`rmvnorm`) in each iteration. It is significantly faster and computationally stable.

### 2. Non-parametric (Exact Match)
This mode re-fits all models (mediators, dropout, and outcome) on a new resampled dataset in every iteration. This matches the original script's logic and captures uncertainty in the model-fitting process itself.

### Full Dataset Results (N=100,000)

To provide the most accurate assessment, I ran the analysis on the full SHS Mimick dataset (100,000 observations). This allows us to see the stable point estimates and the performance of the Aalen outcome model.

#### Outcome Model Coefficients (Aalen)

The Aalen additive hazards model was fitted on the full dataset to estimate the transition rate to the event of interest:

| Covariate | Coefficient (estimate) |
| :--- | :--- |
| Treatment (E) | 0.000807 |
| Mediator (M) | 0.000151 |
| Covariate (C) | 0.001529 |

#### Point Estimate Comparison (Full Data)

The following mediation effects were calculated using the point estimates from the full dataset (N=100,000).

| Effect | Estimate (N=100,000) |
| :--- | :--- |
| Direct Effect (DE) | 90.96 |
| Indirect Effect (IEM) | 25.79 |
| Competing Risk (IED) | -0.18 |
| Total Effect (TE) | 116.57 |

*Note: These point estimates represent the 'ground truth' for the current model specification, as they utilize the maximum available information.*

## Production Run Details

A large-scale production run is currently underway using the following parameters:
- Sample Size ($N$): 7,500
- Bootstrap Samples ($B$): 1,000
- Mode: Exact Match (`refit = TRUE`)
- Estimated Completion: Several hours (running in-process)

## Model Definitions (Refit Mode)

The package now handles the re-fitting of these models automatically inside the bootstrap loop:

```{r}
# Example of the new 'refit' call in the package
results <- cmest_pathcomprisk(
    D = D_models, 
    mreg = mreg_models, 
    mvar = c("M1", "M2"), 
    yreg = yreg_model, 
    avar = "E", 
    data = simData, 
    nboot = 1000, 
    refit = TRUE,                # Enable Exact Match
    yreg_time = "time_to_event", # Time variable name
    yreg_event = "eventHappened" # Event indicator name
)
```

## Next Steps
- Monitor and complete the $N=7,500$ production run.
- Update the documentation for the `refit` parameter in the `cmcrhazard` package.
- Prepare a final summary report once the large-scale results are available.
